### 2014-04-23 article.tex
P1: Start filling in some of the sections, particularly intro/cal.

\documentclass{article}

% Some useful packages.
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{graphicx}
\usepackage{verbatim}

% Reduces margins substantially.
\usepackage{geometry}
\newgeometry{margin=2.5cm}

% Allows headers and footers.
\usepackage{fancyhdr}
\pagestyle{fancy}
% Get rid of annoying line under header.
\renewcommand{\headrulewidth}{0pt}

\lhead{BTTM8 - SURFACE WATER MODELLING - EFFECTS OF DREDGING}
\chead{}
\rhead{}
%\lfoot{}
%\cfoot{}
%\rfoot{}

% Configure References sections. 2 filters are for the 2 sections. (1/practical)
\usepackage[backend=bibtex,style=authoryear,sorting=nyt,dashed=false]{biblatex}
% Changes e.g. \parencite{jones2003} from (Jones 2003) to (Jones, 2003)
\renewcommand*{\nameyeardelim}{\addcomma\space}
\defbibfilter{practical1}{keyword=practical1}
\defbibfilter{practical2}{keyword=practical2}
\bibliography{references/references}

\begin{document}

% Third person throughout. Should mainly be in present tense.

% Practical 1:
\part*{Modelling the effects of dredging on a tidal estuary}

\section{Introduction}

The purpose of this study is to implement a surface water model of the Fowey Estuary. This model is then used to characterise the existing hydrodynamic regime of the Estuary, analyse the sensitivity of the model to changes in the bed friction and evaluate the changes to the regime under a dredging scenario. % drop sensitivity?

The Fowey Estuary is a macrotidal estuary with a mean spring tidal range of \SI{4.8}{\m} and a tidal length of \SI{12}{km} \parencite{uncles2002dependence}. It is located in Cornwall on the south coast of England (see Figure \ref{fig:fowey_estuary_map}). It is fed by two rivers, the Fowey and the Lerryn, with mean discharges of \SI{1.5}{cumecs} and \SI{0.8}{cumecs} respectively. It is tidally dominated, although the contributions of both these rivers was included.

The Estuary is modelled using RMA-2, a two-dimensional depth averaged numerical model that calculates water surface elevation and horizontal velocity components at each timestep \parencite{king1990program}.
The governing equation that is solved is a two-dimensional shallow water equation (2D-SWE). The 2D-SWE is a formulation of the Reynolds averaged Navier-Stokes equation, obtained by integrating it over the vertical coordinate. The 2D-SWE also assumes a hydrostatic pressure distribution, and that the horizontal extent of the model is much larger than the vertical one \parencite{cea2006numerical}.
It uses an implicit finite element numerical integration scheme on an unstructured mesh, and because of the integration scheme used it is numerically stable. Turbulence is modelled using an eddy viscosity approximation. Friction is modelled using the Manning bed friction coefficient.

\begin{figure}[hbp]
    \centering
    \includegraphics[width=9cm]{geogg101_screenshots/practical1/fowey_estuary_map_qgis}
    \caption{Location map of the Fowey Estuary study area, southwest England, UK. }
    \label{fig:fowey_estuary_map}
\end{figure}

\section{Methods}

\subsection{Model}

The model used was two-dimensional depth averaged tidal model named RMA-2 \parencite{king1990program}. It was built within the Surface-water Modelling System (SMS) modelling program, which was used to create the mesh, integrate the bathymetry, apply the river and tidal boundary conditions and control the model runs.

\subsection{Model setup}

\subsubsection{Mesh}
A mesh of the Estuary was created by digitizing the coastline and banks of the rivers. This was resampled at \SI{30}{m} to give an element size of the same order. Beyond the mouth of the Estuary a larger element size of \SI{50}{m} was used. The meshed area, as well as a close-up of a typical section and the transition at the mouth of the Estuary, can be seen in Figure \ref{fig:combined_meshes}. In total, XX \SI{30}{m} elments and YY \SI{50}{m} elements were used. Before running the model, a quality check of the mesh was performed by SMS. There were XX `size changed' warnings, but apart from these the mesh met SMS' quality criteria.

\begin{figure}[hbp]
    \centering
    \includegraphics[width=10cm]{geogg101_screenshots/practical1/combined_meshes}
    \caption{Mesh generated for the Fowey Estuary, with a close-up of two sections. The main mesh shows the boundary conditions for both of the rivers in \SI{}{m.s^{-1}} in red, as well as the initial water level at the sea boundary shown in \SI{}{m} in blue.}
    \label{fig:combined_meshes}
\end{figure}

\subsubsection{Bathymetry}
The bathymetry used was \ldots. The bathymetry was interpolated onto the mesh using \dots.

\subsubsection{Boundary conditions}
At the influx of the Fowey and Lerryn Rivers, mean discharges perpendicular to the boundaries of \SI{1.5}{cumecs} and \SI{0.8}{cumecs} respectively were specified. At the sea boundary, water level was prescribed using an observed time series of tidal levels for the Estuary.

\subsubsection{Run control}
The model was run for \SI{74}{hours} with a \SI{15}{minute} timestep. It was run from a coldstart, with the initial water level at the sea boundary being set to \SI{-0.72}{m}.

\subsection{Calibration}

Throughout the calibration, turbulence was modelled using an eddy viscosity of \SI{9000}{ N.s.m^{-2} }.
Following \textcite{piedra2007residual, sousa2007hydrodynamic}, the  Manning bed friction coefficient was varied to achieve the best fit between the model and the observations. The potential range for the Manning coefficient, $n$, was 0.01 to 0.04. The Nash-Sutcliffe Efficiency (NSE, see \textcite{nash1970river}) and Root Mean Squared Error (RMSE) were calculated for each value of $n$.

\begin{align*}
    RMSE &= \frac{1}{N} \sqrt{ \sum_{t = 1}^N ( O_t - M_t) ^ 2 }\\
    NSE  &= 1 - \frac{\sum_{t = 1}^N ( O_t - M_t) ^ 2 }{\sum_{t = 1}^N ( O_t - O_{mean}) ^ 2 }
\end{align*}

Where $O_t$ is the observed value at timestep $t$, $M_t$ is the measured value, and $N$ is the number of timesteps. The results can be seen if Figure \ref{fig:calibration_stats}. From these four graphs it is clear that the value of $n$ that gives best fit with the observed data is $n = 0.03$, and this value was used for the subsequent models. The two values of $NSE = 0.956, 0.997$ both represent `excellent' fit with data according to the guidelines provided by \textcite{henriksen2008assessment} because they are greater than $0.85$.

\begin{figure}[hbp]
    \centering
    \includegraphics[width=15cm]{geogg101_screenshots/practical1/calibration_stats}
    \caption{RMSE and NSE for velocity magnitude and water level when comparing the simulated and observed values at observation point XXX. The blue dot represents the best fit for each of the four graphs, which is given by a bed friction coefficient of $n = 0.03$.}
    \label{fig:calibration_stats}
\end{figure}

\subsection{Dredging scenario}

To evaluate the effect of dredging on the Estuary, the bathymetry is altered to simulate dredging by \SI{8}{m}. The proposed dredged channel is approximately \SI{8}{m} deeper than the current sea floor near the harbour with a mean difference of \SI{5}{m}, and is \SI{200}{m} wide. This alteration can be seen in Figure \ref{fig:delta_depth}.

\begin{figure}[hbp]
    \centering
    \includegraphics[height=6cm]{geogg101_screenshots/practical1/delta_water_depth_rotated}
    \caption{Difference in bathymetries between undredged and dredged scenarios.}
    \label{fig:delta_depth}
\end{figure}

\newpage
\section{Discussion}

TODO: finish off. Talks about how bathymatric uncertainty can be used to cal model: \textcite{cea2012bathymetric}

\section{Conclusions}

% From handout: bullet points will do here.

\begin{itemize}
    \item The Fowey Estuary is a macrotidal Estuary that is suitable for being modelled by a two-dimensional shallow water model.
    \item It best modelled by RMA-2 when using a Manning bed friction coefficient of 0.03 and an eddy viscosity of \SI{9000}{ N.s.m^{-2} }.
    \item The rising tide produces faster water speeds than the falling tide. During the rising tide, peak water speed is about \SI{8.2}{m/s}.
    \item Dredging the Estuary from the Berth out to sea would reduce the maximum speed by about \SI{0.15}{m.s^{-1}} at the Berth, and \SI{0.10}{m.s^{-1}} at the Inlet.
    \item Dredging the Estuary would have little to no effect further upstream of the dredged channel.
\end{itemize}

\printbibliography[filter=practical1]

\newpage
\setcounter{section}{0}

% Practical 2:
\lhead{BTTM8 - SURFACE WATER MODELLING - EFFECTS OF LAND USE/CLIMATE CHANGE}
\part*{Modelling the effects of land use change and climate change on a catchment area}

\section{Introduction}

The purpose of this study is to calibrate and validate a model of the Karup Catchment in Denmark. This model will then be used to simulate three future emissions scenarios of climate change in 2050: Low, Medium and High.

The Karup Catchment is \SI{440}{km^2}. It is situated in western Denmark and drains into the Limfjord. Geologically, it consists of highly permeable sand and gravel as well as occasionaly lenses of clay. 
The majority of land use is for agriculture (67\%), with forest (18\%), heath (10\%), wetland (4\%) and urban (1\%) making up the remainder \parencite{refsgaard1997parameterisation}.

% TODO: Location map of Karup in Denmark + locations of boreholes/rivers.

MIKE SHE is a deterministic, distributed and physically based modelling system for modelling water flow through catchments \parencite{refsgaard1995mike, refsgaard2010systeme}. It can model precipitation, surface runoff, evapotsranspiration, and saturated zone and unsaturated zone subsurface flow. It can be coupled to MIKE 11, which models one-dimensional water flow through channels by solving the Saint Venant equations \parencite{havno1995mike}.  
It uses a finite difference approach to modelling catchments, and  and can be used with various spatially distributed characteristics such as elevation, land use and soil type. Additionally, the precipitation and evapotsranspiration can be specified for each grid cell in the model \parencite{thompson2012modelling}.

% TODO: Mike she schematic figure (maybe from thompson2004application).

\section{Methodology}

\subsection{Model setup}

\subsubsection{Topology}

\subsubsection{Land cover}

\subsubsection{Unsaturated zone}

\subsubsection{Saturated zone}

\subsubsection{Rivers}

\subsubsection{Precipitation and evapotsranspiration}

\subsubsection{Run control}

\subsection{Calibration}

As can be seen in Table \ref{table:summary_stats}, \dots

MIKE SHE is typically calibrated by modifying the parameters listed in Table \ref{table:calibration_parameters} \parencite{refsgaard1997parameterisation, thompson2004simulation}.

\begin{table}[!h]
    \centering
    \label{table:calibration_parameters}
    \begin{tabular}{l c c}
	Parameter   & Minimum \SI{}{m.s^{-1}} & Maximum \SI{}{m.s^{-1}} \\
	    \hline
	    Horizontal hydraulic condictivity of sand aquifer  & \num{1e-4} & \num{9e-4} \\
	    Vertical hydraulic condictivity of sand aquifer    & \num{1e-5} & \num{9e-5} \\
	    Horizontal hydraulic condictivity of clay lenses   & \num{1e-6} & \num{9e-6} \\
	    Vertical hydraulic condictivity of clay lenses     & \num{1e-7} & \num{9e-7} \\
    \end{tabular}
    \caption{Calibration parameters and the range of appropriate values.}
\end{table}

Using the split-sample method of \textcite{klemevs1986operational}.

\begin{table}[!h]
    \centering
    \label{table:summary_stats}
    \begin{tabular}{l c c c c c c}
	& \multicolumn{3}{c}{Calibration} & \multicolumn{3}{c}{Validation} \\
		    & RMSE & NSE  & R    & RMSE & NSE  & R    \\
	    \hline
	Borehole 5  & 0.19 & 0.63 & 0.82 & 0.23 & 0.52 & 0.77 \\
	Borehole 35 & 0.25 & 0.78 & 0.89 & 0.19 & 0.87 & 0.95 \\
	Borehole 37 & 0.23 & 0.55 & 0.81 & 0.13 & 0.81 & 0.91 \\
	Borehole 65 & 0.26 & 0.69 & 0.84 & 0.18 & 0.90 & 0.95 \\
	    \hline
	Karup       & 0.16 & 0.95 & 0.98 & 0.17 & 0.97 & 0.99 \\
	Hagebro     & 0.06 & 0.92 & 0.97 & 0.07 & 0.92 & 0.97 \\

    \end{tabular}
    \caption{Summary statistics comparing the observed and simulated values for the four boreholes and two rivers. The calibration period was from 1971 to 1975 and the validation period was from 1976 to 1981.}
\end{table}

\subsection{Scenarios}

\subsubsection{Changing land use}

\subsubsection{Climate change}

\section{Results}

\subsection{Effects of land use change}

\subsection{Effects of climate change}

\section{Conclusions}
Conclusions to go here\ldots

\printbibliography[filter=practical2]

\end{document}